### 2021-01-28 ch01_introduction/introduction.tex
Added correct line breaks
Upadted author contributions

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). Unless
stated otherwise, results, discussions and conclusions presented from these
studies originate from the author of this thesis. A detailed list of
corresponding references is given at the beginning of the relevant chapters.

>>>>

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). Wherever
material from these studies is presented in this thesis, the pronoun
\enquote{we} is used to increase readability by avoiding the passive voice and
to acknowledge all involved contributors. However, unless stated otherwise, all
contents from these publications (text, figures and tables) shown in this
thesis originate from the author of this thesis. A detailed list of
contributions to these studies (including their full reference) is given in the
corresponding chapters.

### 2021-01-28 ch04_papers_ecs_tcr_assessment/papers_ecs_tcr_assessment.tex
Upadted author contributions

In order to reduce uncertainties in multi-model climate projections, a first
important step is the assessment of the desired target variables in the
corresponding climate model ensemble. This is particularly relevant when a new
generation of \acp{ESM} is published that includes considerable modifications
compared to the respective predecessor model versions. Currently, we are in a
situation like this with the \acs{CMIP}6 ensemble, for which new data is still
released every day (as of \TheMonth{}). In light of the upcoming international
climate assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of
policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6
models is crucial since they provide vital information about the future climate
of planet Earth. As the \ac{ESMValTool} is an excellent tool that allows for a
quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we use it to assess
and analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to
corresponding results of predecessor model generations and international
climate assessments. This work is already published in two scientific papers:
\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
author of this thesis contributed their figure 8 (a bar chart showing the
\ac{ECS} for the three climate model generations \acs{CMIP}3, \acs{CMIP}5 and
\acs{CMIP}6), their figure 10 (map plots showing cloud-related feedback
parameters for the \acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
\cref{fig:04:cloud_feedback_parameters}) and discussion about \ac{ECS} of the
\acs{CMIP}6 models. For \textcite{Meehl2020}, the author of this thesis
contributed all figures and tables of the study. This chapter first presents
the evaluation of climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6
ensembles (\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding
\ac{ECS} and \ac{TCR} values into historical context
(\cref{sec:04:historical_context}). Finally, possible reasons for the apparent
increase of climate sensitivity in the \acs{CMIP}6 models are discussed in
detail (\cref{sec:04:possible_reasons_high_ecs_cmip6}).

>>>>

In order to reduce uncertainties in multi-model climate projections, a first
important step is the assessment of the desired target variables in the
corresponding climate model ensemble. This is particularly relevant when a new
generation of \acp{ESM} is published that includes considerable modifications
compared to the respective predecessor model versions. Currently, we are in a
situation like this with the \acs{CMIP}6 ensemble, for which new data is still
released every day (as of \TheMonth{}). In light of the upcoming international
climate assessment of the \ac{IPCC}'s \acs{AR}6, the evaluation of
policy-relevant metrics like \ac{ECS} or \ac{TCR} for these new \acs{CMIP}6
models is crucial since they provide vital information about the future climate
of planet Earth. As the \ac{ESMValTool} is an excellent tool that allows for a
quick and robust evaluation of new \ac{CMIP} data as soon as it gets published
on the \ac{ESGF} servers (see \cref{sec:03:esmvaltool}), we use it to assess
and analyze \ac{ECS} and \ac{TCR} of the \acs{CMIP}6 models and compare it to
corresponding results of predecessor model generations and international
climate assessments. This work is already published in two scientific papers:
\textcite{Bock2020} and \textcite{Meehl2020}. For \textcite{Bock2020}, the
author of this thesis contributed their figure 8 (a bar chart showing the
\ac{ECS} for several \acs{CMIP}3, \acs{CMIP}5 and \acs{CMIP}6 models), their
figure 10 (map plots showing cloud-related feedback parameters for the
\acs{CMIP}5 and \acs{CMIP}6 \acp{MMM}; see
\cref{fig:04:cloud_feedback_parameters}), code to create these figures with the
\ac{ESMValTool} and text for the manuscript (in particular a section about the
evaluation of \ac{ECS}). For \textcite{Meehl2020}, the author of this thesis
contributed all figures and tables, code to produce these with the
\ac{ESMValTool} and text for the manuscript. This chapter first presents the
evaluation of climate sensitivity for the \acs{CMIP}5 and \acs{CMIP}6 ensembles
(\cref{sec:04:evaluation_ecs_and_tcr}) and puts the corresponding \ac{ECS} and
\ac{TCR} values into historical context (\cref{sec:04:historical_context}).
Finally, possible reasons for the apparent increase of climate sensitivity in
the \acs{CMIP}6 models are discussed in detail
(\cref{sec:04:possible_reasons_high_ecs_cmip6}).

### 2021-01-28 ch05_paper_ecs/paper_ecs.tex
Adapted copyright statements
Upadted author contributions

One of the most important sources of information for this purpose is
undoubtedly observational data. As presented in
\cref{subsec:02:emergent_constraints}, emergent constraints are a beneficial
tool to assess the accuracy of climate model projections and reduce associated
uncertainties with appropriate observations of the Earth system. For \ac{ECS},
a large variety of different emergent constraints has been developed over the
years, which we now evaluate on the new \acs{CMIP}6 ensemble. In total we
assess eleven emergent constraints (selected by their availability in the
\ac{ESMValTool}), out of which the most are related to cloud feedbacks, since
these constitute the most important source of uncertainty for \ac{ECS}
\commentcite{Boucher2013, Flato2013}{see also
  \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
constraints have been derived on the \acs{CMIP}3 or \ac{CMIP}5 ensemble, the
\ac{CMIP}6 models offer an exciting possibility to perform out-of-sample
testing using the previously unseen data. We quantify the skill of the emergent
constraints with three metrics: the coefficient of determination $R^2$ of the
linear relationship, the statistical significance using a two-sided $t$-test
based on the Pearson correlation coefficient $r$ (see \cref{eq:02:t}) and the
constrained range of \ac{ECS} based on the constrained \ac{PDF} (see
\cref{eq:02:pdf_y_given_x0}). This analysis has already been published in
\textcite{Schlund2020a}. In this chapter, we first introduce the eleven
emergent constraints on \acs{ECS} and evaluate them on \acs{CMIP}5 and
\acs{CMIP}6 data (\cref{sec:05:comparison_of_emergent_constraints}). After
that, we provide a discussion of the results and draw important conclusions
from them (\cref{sec:05:discussion_and_conclusions}).

@@

\begin{figure}[p]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/scatterplot_merged_training_data_BRI.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/target_distribution_training_data_BRI.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/scatterplot_merged_training_data_COX.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
  ch05_paper_ecs/figs/target_distribution_training_data_COX.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:d}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/scatterplot_merged_training_data_LIP.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:e}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
  ch05_paper_ecs/figs/target_distribution_training_data_LIP.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:f}
  \end{subfigure}
  \caption{xD. Updated from
    \textcite{Schlund2020a}.}
  \label{fig:05:bri_cox_lip}
\end{figure}

>>>>

One of the most important sources of information for this purpose is
undoubtedly observational data. As presented in
\cref{subsec:02:emergent_constraints}, emergent constraints are a beneficial
tool to assess the accuracy of climate model projections and reduce associated
uncertainties with appropriate observations of the Earth system. For \ac{ECS},
a large variety of different emergent constraints has been developed over the
years, which we now evaluate on the new \acs{CMIP}6 ensemble. In total we
assess eleven emergent constraints (selected by their availability in the
\ac{ESMValTool}), out of which the most are related to cloud feedbacks, since
these constitute the most important source of uncertainty for \ac{ECS}
\commentcite{Boucher2013, Flato2013}{see also
  \cref{subsec:02:climate_feedbacks}}. Since all of the evaluated emergent
constraints have been derived on the \acs{CMIP}3 or \ac{CMIP}5 ensemble, the
\ac{CMIP}6 models offer an exciting possibility to perform out-of-sample
testing using the previously unseen data. We quantify the skill of the emergent
constraints with three metrics: the coefficient of determination $R^2$ of the
linear relationship, the statistical significance using a two-sided $t$-test
based on the Pearson correlation coefficient $r$ (see \cref{eq:02:t}) and the
constrained range of \ac{ECS} based on the constrained \ac{PDF} (see
\cref{eq:02:pdf_y_given_x0}). This analysis has already been published in
\textcite{Schlund2020a}. For this paper, the author of this thesis lead the
writing and the analysis and contributed the code to reproduce this analysis
including all figures and tables with the \ac{ESMValTool}. In this chapter, we
first introduce the eleven emergent constraints on \acs{ECS} and evaluate them
on \acs{CMIP}5 and \acs{CMIP}6 data
(\cref{sec:05:comparison_of_emergent_constraints}). After that, we provide a
discussion of the results and draw important conclusions from them
(\cref{sec:05:discussion_and_conclusions}).

@@

\begin{figure}[p]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/scatterplot_merged_training_data_BRI.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/target_distribution_training_data_BRI.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/scatterplot_merged_training_data_COX.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
  ch05_paper_ecs/figs/target_distribution_training_data_COX.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:d}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
      ch05_paper_ecs/figs/scatterplot_merged_training_data_LIP.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:e}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{
  ch05_paper_ecs/figs/target_distribution_training_data_LIP.pdf}
    \caption{}
    \label{fig:05:bri_cox_lip:f}
  \end{subfigure}
  \caption{xD. \AdaptedFrom{Schlund2020a}.}
  \label{fig:05:bri_cox_lip}
\end{figure}
